---
title: "ARE213 PSet1"
output: html_document
date: "2023-09-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

0. Load packages
```{r}
library(dplyr)
library(haven)
library(tidyr)
```
0. Load data
```{r}
#on PC
my_data<-read_dta("C:/Users/Ken/Documents/Github/HP-ARE213-Psets/Pset 1/pset1.dta")

#On Mac
#my_data<-read_dta("/Users/alexadia/Documents/GitHub/HP-ARE213-Psets/Pset 1/pset1.dta")
```
1a. Clean the missing values from cardiac to wgain. For cardiac, lung, diabetes, herpes, hydra, hemo, chyper, phyper, eclamp, incervix, pre4000, preterm, missing is encoded by 8 (not on certificate) and 9 (not classifiable). For tobacco and alcohol, missing is encoded by 9 (unknown). For cigar6, missing is given by 6 (unknown or not stated). For drink5, missing is given by 5 (unknown or not stated). For wgain, missing is given by 99 (unknown or not stated).
```{r}
#I need to tabulate these all still lol

#For variables that can take on 8 and 9
cleanset1 <- c("cardiac", "lung", "diabetes", "herpes", "chyper", "phyper", "pre4000", "preterm")

my_data <- my_data %>%
  mutate(across(all_of(cleanset1), ~ifelse(. %in% c(8, 9), NA, .)))
#check cardiac
unique(my_data$cardiac)

#For variables that only take on 9 as missing
cleanset2 <- c("tobacco", "alcohol")
my_data <- my_data %>%
  mutate(across(all_of(cleanset2), ~na_if(., 9)))
#check tobacco
unique(my_data$tobacco)

#For drink5
my_data <- my_data %>%
  mutate(drink5 = ifelse(drink5 == 5, NA, drink5))
#check drink5
unique(my_data$drink5)

#For cigar6
my_data <- my_data %>%
  mutate(cigar6 = ifelse(cigar6 == 6, NA, cigar6))
#check drink5
unique(my_data$cigar6)

#For wgain
my_data <- my_data %>%
  mutate(wgain = ifelse(wgain == 99, NA, wgain))
#check wgain
unique(my_data$wgain)
```
1b.Recode all indicator variables. We see that rectype, pldel3, dmar, csex, anemia, cardiac, lung, diabetes, herpes, chyper, phyper, pre4000, preterm, tobacco, and alcohol are indicators.
```{r}
#Check which variables take on the form 1, 2, or missing
sapply(my_data, function(x) all(x %in% c(1, 2, NA)))
#Turn each indicator variable into the form 0 v 1 by mutating such that 2's become 0's.
indicator_set <- c("rectype", "pldel3", "dmar", "csex", "anemia", "cardiac", "lung", "diabetes", "herpes", "chyper", "phyper", "pre4000", "preterm", "tobacco", "alcohol")
for(var in indicator_set){
my_data<-my_data %>%
  mutate_at(var, funs(ifelse(.==2, 0, .)))
}
#Check if this works - everything should return false using our original check
sapply(my_data, function(x) all(x %in% c(1, 2, NA)))

```
Next we recode the other variables as directed. 
```{r}
#Recode mrace3 as categorical variable (so a factor variable)
class(my_data$mrace3) #starts off as numeric
my_data<-my_data%>%mutate(mrace3 = factor(mrace3))
#Check
class(my_data$mrace3) #confirmed as factor

#Coarsen ormoth and orfath into indicator variables. 
unique(my_data$ormoth)
unique(my_data$orfath)
my_data<-my_data%>% mutate(ormoth = ifelse(ormoth > 1, 1, ormoth),
                           orfath = ifelse(orfath > 1, 1, orfath))
#Check
unique(my_data$ormoth)
unique(my_data$orfath)

#Drop stresfip, birmon, and weekday
unique(my_data$stresfip) #Could convert this into a regional variable
unique(my_data$birmon) #Could turn this into a quarter?
unique(my_data$weekday) #could treat like a weekday vs weekend binary variable
#Now I drop them.
my_data<-my_data%>% select(-c(stresfip, birmon, weekday))
```
1c. Produce the analysis data set. It has the desired set of variables.
The naniar library has a neat little MCAR test function that runs Little's test whether data is missing completely at random. It appears like it fulfills the MCAR assumption. 
```{r}
library(naniar)
mcar_test(my_data)
```
Then I drop all things
```{r}
my_data<-drop_na(my_data)
```
1d. 
I use the stargazer function to create an overall table, then dplyr to get tobacco and non-tobacco datasets to get different spltis
```{r}
library(stargazer)
stargazer(as.data.frame(my_data), type = "text", summary = TRUE)
#We then split out the data
tobacco_data<-my_data%>%filter(tobacco==1)%>%select(-tobacco)
stargazer(as.data.frame(tobacco_data), type = "text", summary = TRUE)
notobacco_data<-my_data%>%filter(tobacco==0)%>%select(-tobacco)
stargazer(as.data.frame(notobacco_data), type = "text", summary = TRUE)
```

Q2. Our goal will be to analyze the causal effects of maternal smoking during pregnancy on infant birth weight.
a. Compute the mean difference in birthweight in grams by smoking status. Is this difference likely to be causal? Provide some evidence for or against.
```{r}
#One way to do this is via the aggregate() function:
mean_diff<-my_data%>%aggregate(dbrwt~tobacco, FUN = mean)
mean_diff

#If we want to t-test, we can split it out into two different datasets then use the t.test command
nonsmokers<-my_data%>%filter(tobacco==0)%>%select(dbrwt)
smokers<-my_data%>%filter(tobacco==1)%>%select(dbrwt)
dbrwt_test<-t.test(nonsmokers, smokers)
dbrwt_test
```
There is a statistically significant difference across smoking vs nonsmoking mothers in terms of mean birthweight. This is unlikely to be causal given the presence of several different confounders in the dataset. For example, an impoverished mother may smoke more as a stress response and may be unable to invest time and money into practices and resources that would yield a healthier birth - this is one such example in which this mean estimate would not be causal. 

b. Classify the variables in the dataset into different types depending on how they are potentially related to the treatment and the outcome; justify your choices and be explicit about the assumptions you are making.4 Use your classification to decide on the list of covariates you’ll keep for covariate adjustment for the rest of the problem set.
```{r}
#Based on the above, we can generate a new object that has our outcome and all covariates
test_data<-my_data%>%select(dbrwt, tobacco, ormoth, mrace3, dmage, dmeduc, dmar,adequacy, monpre, nprevist, isllb10, csex, dplural, anemia, cardiac, lung, diabetes, herpes, chyper, phyper)
```

3. We will now investigate different methods based for covariate adjustment, starting from regression.
a. Use a basic, uninteracted linear regression model to estimate the impact of smoking and report your estimates. Under what circumstances does it identify the average treatment effect (ATE)?
```{r}
model3a<-lm(dbrwt ~ tobacco + ormoth + mrace3 + dmage + dmeduc + dmar + adequacy + monpre + nprevist + isllb10 + csex + dplural + anemia + cardiac + lung + diabetes + herpes + chyper + phyper, test_data)
model3a
```
b. Is the estimate in the previous question sensitive to dropping controls one at a time? What do you learn from this exercise?
```{r}
#without phyper
model3b1<-lm(dbrwt ~ tobacco + ormoth + mrace3 + dmage + dmeduc + dmar + adequacy + monpre + nprevist + isllb10 + csex + dplural + anemia + cardiac + lung + diabetes + herpes + chyper, test_data)
#without phyper and dmar
model3b2<-lm(dbrwt ~ tobacco + ormoth + mrace3 + dmage + dmeduc + adequacy + monpre + nprevist + isllb10 + csex + dplural + anemia + cardiac + lung + diabetes + herpes + chyper, test_data)
#without phyper, dmar, and adequacy
model3b3<-lm(dbrwt ~ tobacco + ormoth + mrace3 + dmage + dmeduc + monpre + nprevist + isllb10 + csex + dplural + anemia + cardiac + lung + diabetes + herpes + chyper, test_data)
#without phyper, dmar, adequacy, and mrace3
model3b4<-lm(dbrwt ~ tobacco + ormoth + dmage + dmeduc + monpre + nprevist + isllb10 + csex + dplural + anemia + cardiac + lung + diabetes + herpes + chyper, test_data)
#without phyper, dmar, adequacy, mrace3, and nprevisit
model3b5<-lm(dbrwt ~ tobacco + ormoth + dmage + dmeduc + monpre + isllb10 + csex + dplural + anemia + cardiac + lung + diabetes + herpes + chyper, test_data)
#without phyper, dmar, adequacy, mrace3, nprevisit, and isllb10
model3b5<-lm(dbrwt ~ tobacco + ormoth + dmage + dmeduc + monpre + csex + dplural + anemia + cardiac + lung + diabetes + herpes + chyper, test_data)
```
c. For this part only, extend the OLS specification from question 3(a) to control for the covariates using a more flexible functional form. Describe the specification you picked. What are the potential benefits and drawbacks of this approach?
```{r}
#With interactions across all other covariates
model3c <- lm(dbrwt ~ (tobacco + ormoth + mrace3 + dmage + dmeduc + dmar + adequacy + monpre + nprevist + isllb10 + csex + dplural + anemia + cardiac + lung + diabetes + herpes + chyper + phyper)^2, data = test_data)
model3c

#with interactions across only tobacco (what I saw in Jackee's code)
model3c <- lm(dbrwt ~ tobacco + ormoth + mrace3 + dmage + dmeduc + dmar + adequacy + monpre + nprevist + isllb10 + csex + dplural + anemia + cardiac + lung + diabetes + herpes + chyper + phyper + tobacco:ormoth + tobacco:mrace3 + tobacco:dmage + tobacco:dmeduc + tobacco:dmar + tobacco:adequacy + tobacco:monpre + tobacco:nprevist + tobacco:isllb10 + tobacco:csex + tobacco:dplural + tobacco:anemia + tobacco:cardiac + tobacco:lung + tobacco:diabetes + tobacco:herpes + tobacco:chyper + tobacco:phyper, data = test_data)
model3c

```
d.For this part only, add to the specification of question 3(a) some “bad controls.”
Check if your estimate changes and discuss the direction of the change.
```{r}
#Re-select test data to include things we want 
bad_control_data<-my_data%>%select(dbrwt, tobacco, ormoth, mrace3, dmage, dmeduc, dmar,adequacy, monpre, nprevist, isllb10, csex, dplural, anemia, cardiac, lung, diabetes, herpes, chyper, phyper, dgestat)

model3d<-lm(dbrwt ~ tobacco + ormoth + mrace3 + dmage + dmeduc + dmar + adequacy + monpre + nprevist + isllb10 + csex + dplural + anemia + cardiac + lung + diabetes + herpes + chyper + phyper+dgestat, bad_control_data)
```
e. Produce the Oaxaca-Blinder estimator for the ATE and ATT. Describe the exact steps you have used. Does your answer differ substantially from the one in 3(a)? Discuss. (AA Note: unclear if proper method - group by tobacco with the oaxaca package?)
```{r}
#First split the dataset by the exposure (tobacco)
nonsmokers<-test_data%>%filter(tobacco==0)
smokers<-test_data%>%filter(tobacco==1)

#Next get the model results for each
nonsmokers_model<-lm(dbrwt~., nonsmokers)
smokers_model<-lm(dbrwt~., smokers)

#Predict each
nonsmokers_predict<-predict(nonsmokers_model)
smokers_predict<-predict(smokers_model)

#First get ATE using these predicted values
oaxaca_ATE <- mean(smokers_predict) - mean(nonsmokers_predict)

#Next get ATT
oaxaca_ATT <- mean(smokers$dbrwt - smokers_predict)
```
4. Next on the list is the propensity score approach.
a.Estimate the propensity score using a logit specification without nonlinear terms and interactions. Discuss which covariates appear most predictive of maternal smoking and whether this matches your expectation.
```{r}
#Run logit specification
pscore_model<-glm(tobacco~ormoth + mrace3 + dmage + dmeduc + dmar + adequacy + monpre + nprevist + isllb10 + csex + dplural + anemia + cardiac + lung + diabetes + herpes + chyper + phyper, family = binomial(link = "logit"), test_data)
#predict outputs from the model to get the propensity score
pscore<-predict(pscore_model, type="response")
```
b. How good is the overlap of propensity scores between the treated and untreated groups?
```{r}
#Separate out the propensity 
smoker_propensity_scores <- pscore[test_data$tobacco == 1]
nonsmoker_propensity_scores <- pscore[test_data$tobacco == 0]

#Get plot for overlap
library(ggplot2)

#For the plot, get everything a factor variable for the pscores by each treatment arm
treatment_factor <- factor(rep(c("Smokers", "Nonsmokers"), times = c(length(smoker_propensity_scores), length(nonsmoker_propensity_scores))))

plot_data <- data.frame(
  pscores = c(smoker_propensity_scores, nonsmoker_propensity_scores),
  treatment = treatment_factor
)

#Get plots that overlap
q4bplot<-ggplot(plot_data, aes(x = pscores, fill = treatment)) +
  geom_histogram(alpha = 0.5) +
  labs(x = "propensity scores", y = "density") +
  scale_fill_manual(values = c("red", "blue")) +
  theme_minimal()

q4bplot
```
c.Assess whether this logit specification has been sufficient to balance the covariates.
```{r}

```
d. Estimate the ATE and ATT via propensity score blocking or matching
```{r}
#load the MatchIt package
library(MatchIt)

#get a dataset to identify match
match_data<-data.frame(smoking=test_data$tobacco, pscores=pscore, birthweight=test_data$dbrwt)

#next run the match via the matchit command- picking nearest neighbor
match_data<-matchit(smoking ~ pscores, data=match_data, method="nearest")
match_data<-match.data(match_data) #syntax is clumsy here - sorry!

#next get the ATE and ATT
match_treated <- subset(match_data, smoking == 1)
match_untreated <- subset(match_data, smoking == 0)

match_ATE <- mean(match_treated$birthweight) - mean(match_untreated$birthweight)
match_ATT <- mean(match_treated$birthweight-match_untreated$birthweight)
```
e. Estimate the ATE and ATT via propensity score reweighting. Include the formulas you used.
```{r}
#Use the WeightIt package
library(WeightIt)

#First ATE
weight_data<-weightit(tobacco~ormoth + mrace3 + dmage + dmeduc + dmar + adequacy + monpre + nprevist + isllb10 + csex + dplural + anemia + cardiac + lung + diabetes + herpes + chyper + phyper, test_data, method="ps", estimand="ATE")
#reweighting
weights<-weights(weight_data)
#we add the weighted data as a column to test_data
weighted_data<-test_data%>%mutate(weights=weights)
#get ATE
reweight_ATE <- weighted.mean(weighted_data$dbrwt[weighted_data$tobacco == 1], w = weights[weighted_data$tobacco == 1])- weighted.mean(weighted_data$dbrwt[weighted_data$tobacco == 0], w = weights[weighted_data$tobacco == 0])

#Next get ATT
weight_data<-weightit(tobacco~ormoth + mrace3 + dmage + dmeduc + dmar + adequacy + monpre + nprevist + isllb10 + csex + dplural + anemia + cardiac + lung + diabetes + herpes + chyper + phyper, test_data, method="ps", estimand="ATT")
#reweighting
weights<-weights(weight_data)
#we add the weighted data as a column to test_data
weighted_data<-test_data%>%mutate(weights=weights)
#flag observations by smoker/nonsmoker
smoker_flag <- ifelse(weighted_data$tobacco == 1, 1, 0)
#split data out into weighted data by smokers and nonsmokers
weighted_dbrwt_smokers <- sum(weighted_data$dbrwt[smoker_flag == 1] * weights[smoker_flag == 1]) / sum(weights[smoker_flag == 1])
weighted_dbrwt_nonsmokers <- sum(weighted_data$dbrwt[smoker_flag == 0] * weights[smoker_flag == 0]) / sum(weights[smoker_flag == 0])
#get ATT
reweight_ATT <- mean(weighted_dbrwt_smokers - weighted_dbrwt_nonsmokers)
```
5. Finally, try doubly-robust methods
a. Estimate the ATE and ATT using one the “mixed methods”: regression adjustment combined with propensity score blocking, matching, or reweighting.
```{r}
#re-run the pscore match
pscore_model<-glm(tobacco~ormoth + mrace3 + dmage + dmeduc + dmar + adequacy + monpre + nprevist + isllb10 + csex + dplural + anemia + cardiac + lung + diabetes + herpes + chyper + phyper, family = binomial(link = "logit"), test_data)
pscore<-predict(pscore_model, type="response")

#use matchit to get matched
match_data<-matchit(tobacco ~ pscore, data=test_data, method="nearest")
match_data<-match.data(match_data) #syntax is clumsy here - sorry!

#next rerun regression
doublyrobust_model<-lm(dbrwt~tobacco + ormoth + mrace3 + dmage + dmeduc + dmar + adequacy + monpre + nprevist + isllb10 + csex + dplural + anemia + cardiac + lung + diabetes + herpes + chyper + phyper, match_data)
doublyrobust_coefs<-coef(doublyrobust_model)

#doubly robust ATE
doublyrobust_ATE <- doublyrobust_coefs["tobacco"]
doublyrobust_ATT<- mean(match_data$dbrwt[match_data$tobacco == 1]) - mean(match_data$dbrwt[match_data$tobacco == 0])
```
b. Assuming constant effects, estimate the causal effect using post-double selection LASSO. Use flexible specifications with some polynomials and interaction terms. (Defining these nonlinear terms — “feature engineering” — for the use in LASSO may require some thinking.) How many covariates did you start with? How many were selected by LASSO in the outcome regression? How many in the propensity score regression? How many overlapped? In this application, would you get a very different answer if you didn’t include the covariates chosen in the propensity score regression?
```{r}
#load glmnet and caret
library(caret)
library(glmnet)

#Split into outcome and exposure
q5y<-test_data%>%select(dbrwt)
q5d<-test_data%>%select(tobacco)
q5x<-test_data%>%select(-dbrwt, -tobacco)

#If I wanted to have interactions between all my terms
#q5x<-model.matrix(~ (tobacco + ormoth + mrace3 + dmage + dmeduc + dmar + adequacy + monpre + nprevist + isllb10 + csex + dplural + anemia + cardiac + lung + diabetes + herpes + chyper + phyper)^2, data = q5x)
#Because my computer would die, I do the following instead
for (covariate in colnames(q5x)) {
  if (covariate != "nprevist") {
    interaction_term <- q5x$nprevist * q5x[[covariate]]
    q5x[[paste("nprevist", covariate, sep = "")]] <- interaction_term
  }
}
#Add a polynomial for age and nprevisit
q5x<-q5x%>%mutate(dmage2=dmage*dmage,
                  nprevisit2=nprevist*nprevist)

#get everything as a numeric so I can apply scale
q5x <- q5x[, sapply(q5x, is.numeric)]

#Standardize data before double selection
q5x <- scale(q5x)

#convert d and y into a numeric vs a one column dataframe
q5y <- q5y[[1]]
q5d <- q5d[[1]]

#LASSO Y on X and get relevant coefficients
lasso1<-cv.glmnet(q5x, q5y, alpha=1)
best_lambda1<-lasso1$lambda.min
lasso_coef1<-coef(lasso1, s=best_lambda1)

#LASSO D on X and get relevant coefficients
lasso2<-cv.glmnet(q5x, q5d, alpha=1)
best_lambda2<-lasso2$lambda.min
lasso_coef2<-coef(lasso2, s=best_lambda2)

#Get the covariates that are overlapping
union_covariates <- intersect(
  rownames(lasso_coef1)[-1], 
  rownames(lasso_coef2)[-1]  
)
select_x<- q5x[, union_covariates]

#Get everything in a dataset
postlasso_data<-data.frame(q5y, q5d, select_x)
postlasso_model<-lm(q5y~., postlasso_data)

#Assume this is the ATE?
lasso_ATE <- coef(postlasso_model)["q5d"]
```

6. Compare the estimates and standard errors between regression adjustment methods, propensity score methods, and doubly robust methods you have used. Concisely and coherently summarize your results above providing some intuition. Write it like you would the conclusion of a paper. In this summary, describe whether you think your best estimate of the effects of smoking is credibly identified; state why or why not.
```{r}

```